Limited plasticity but increased variance in physiological rates across ectotherm populations under climate change

### 2025-01-02 ms.qmd
Adjusted qmd

To explore the potential consequences of the impacts that human-induced climate change may have on variance in physiological rates we fit a model that included an interaction between acclimation type, habitat type, latitude and longitude (referred to as “Model 6”). We assumed that any change in $lnVR_{Q_{10}}$ across latitude and longitude could vary by habitat type (i.e., an interaction between habitat). We used non-linear tensors for latitude and longitude as any response could be complicated by local factors (e.g., altitude). Our model included random effects of species, trait, phylogeny and study. We predicted the expected change in $lnVR_{Q_{10}}$ for each wild population in our dataset at its specific populations latitude and longitude. We first converted the predicted $lnVR_{Q_{10}}$ to a 1°C change as opposed to 10°C to better map to relevant changes in temperature coinciding with climate change:

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "Meta-analysis results for different habitats. In both panels, thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets and raw effects are weighted by their precision (inverse sampling variance). $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. (A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. Overall mean physiological rates ($\\mu$) across the habitats are provided in the results for simplicity and only contrasts between acute and acclimation $lnRR_{Q_{10}}$ are shown (B) Mean acute and acclimation $lnVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B.  For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false

>>>>

To explore the potential consequences of the impacts that human-induced climate change may have on variance in physiological rates we fit a model that included a non-linear smoother between latitude and longitude and an interaction between effect type and habitat type while correcting for acclimation time and maximum temperature (referred to as “Model 6”). We assumed that any change in $lnVR_{Q_{10}}$ across latitude and longitude could vary by habitat type (i.e., an interaction between habitat). We used non-linear tensors for latitude and longitude as any response could be complicated by local factors (e.g., altitude). Our model included random effects of species, trait, phylogeny and study. We predicted the expected change in $lnVR_{Q_{10}}$ for each wild population in our dataset at its specific populations latitude and longitude. We first converted the predicted $lnVR_{Q_{10}}$ to a 1°C change as opposed to 10°C to better map to relevant changes in temperature coinciding with climate change:

@@

```{r, fig-2}
#| label: fig-2
#| fig-cap: "Meta-analysis results for different habitats. In both panels, thick bars are 95% confidence intervals (CI) and thin bars 95% prediction intervals (PI). $\\beta$ values are the contrasts between acute and acclimation means within each habitat. $\\mu$ values are the overall meta-analytic means averaged across acute and acclimation types within each habitat type. In both cases, their 95% CI's are indicated within square brackets and raw effects are weighted by their precision (inverse sampling variance). $p_{MCMC}$ values are the posterior probability of the contrast or overall meta-analytic mean being different from zero. (A) Mean acute and acclimation $lnRR_{Q_{10}}$ across ectotherms in marine, freshwater, and terrestrial habitats. Overall mean physiological rates ($\\mu$) across the habitats are provided in the results for simplicity and only contrasts between acute and acclimation $lnRR_{Q_{10}}$ are shown. Percentages refer to the percentage change in physiological rates between acclimation and acute $lnRR_{Q_{10}}$ (B) Mean acute and acclimation $lnVR_{Q_{10}}$ across ectotherms in marine, freshwater and terrestrial habitats. Percentages refer to the percentage change in physiological rate variance for a 10°C temperature change. For both plots, k = total number of effect size estimates while the numbers in brackets indicate the number of species. Sample sizes are the same for panel A and B.  For ease of visualisation, all the raw data plotted for both acute and acclimation type effect sizes are presented as circles."
#| warning: false
#| echo: false
